Toward a Comparable Corpora of Latvian, Russian and English Tweets

### 2017-04-18 paper.tex
Sascha changes.

\begin{abstract} 
A possibility of building a trilingual Latvian-Russian-English corpus of tweets from Riga, Latvia is investigated. Such a corpus, when constructed, might be of great use for training machine translation models, examining cross-lingual phenomena and studying the population of Riga. This pilot study shows that it is feasible to build such a resource. The resulted corpus is made publicaly available and might be used to construct a large comparable corpus.
\end{abstract}

@@

Comparable corpora are widely used by the natural language processing community to build machine translation or information retrieval models. The goal of this work is to investigate whether it is possible to build a comparable linguistic resource that originates from Riga, Latvia. Riga is a great location because it is a multilingual city where Latvian and Russian are widely used in everyday life, and English is spread in tourism and commerce. Being linked to a location, the resulting corpus might benefit not only the language engineers and linguists, but also social scientists.

@@

The analysis of the tweets, shows that they reflect real world events, such as national celebration or international political affair. While all three languages are represented---45.5\% tweets are in Latvian, 33.9\% in Russian and 20.7\% in English---not all events are equally discussed in all languages (Section~\ref{sec:timeline}). By studying users' tweeting habits, we see that the majority of users (83.3\%) mostly tweets in one language (Section~\ref{sec:lang-use}).

The properties of the corpus correspond to the expectation to reflect the real world events and language use proportion, but its size is to small to draw solid conclusions. However, the construction of a reliable corpus is the matter of the data collection procedure, because, as this study shows, there is relevant data on Twitter.

The resulted dataset is distributed as a supplement to the paper and available online at \texttt{http://...}

@@

251\,083 tweets were collected from the period from the 1st of November 2016 to the 31st of March 2017. On April 14th 2017, the collection was rehydrated by querying the Twitter API with the collected tweet IDs to get rid of the deleted tweets. In addition, the tweets that originated retweets were included to the collection: the JSON representation of a retweet includes the original tweet, which was extracted and added to the collection. Rehydrated and expanded collection resulted to 220\,883 tweets. Table~\ref{tab:tweet-counts} summarises the number of collected tweets.

@@

Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare.\footnote{\url{https://foursquare.com}} This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' most of the time follow a predefined template and thus do not reflect real language use.

@@

Table~\ref{tab:client-counts} shows the top ten most popular clients in the rehydrated collection. Together with the tweets originated from Foursquare, tweets from Instagram,\footnote{\url{https://www.instagram.com}} an image sharing service, and Endomondo,\footnote{\url{https://www.endomondo.com}} a workout tracking service, were removed. Tweets written using the ``Word Cities'' client that posts weather reports and the Linkis client---a promotion website---were also removed.

The final collection resulted into 136\,067 tweets which are in Latvian, Russian or English and created after the 1st of November 2016. The language of a tweet is provided by the corresponding field in the tweet JSON representation.

@@

Figure~\ref{fig:timeline} shows the bandwidth of tweets in time for all three languages. There are several peaks in Twitter usage. Some of them affect all three languages, as in early January, some of them affect only one language, as in late January.

If the Twitter behaviour is affected by the events in the real world, then the peaks should correspond to events in the real word. The difference in peaks, could then be explained as there are different real word events that trigger discussions on Twitter in Latvian, Russian and English. Table~\ref{tab:timeline-lang-corr} suggests, that tweets in Latvian and English share similar behaviour. The Russian tweet timeline is distinct from both timelines, tough its behaviour is more similar to the Latvian timeline than to English.

@@

11th of November is \href{https://en.wikipedia.org/wiki/L\%C4\%81\%C4\%8Dpl\%C4\%93sis_Day}{L\={a}\v{c}pl\={e}sis Day}, a memorial day for soldiers who fought for the independence of Latvia. 18 November is the Proclamation Day of the Republic of Latvia. Numerous events take place across the country including candle placing on and by the wall of Riga Castle, torchlight processions and fireworks. The volume of tweets peaks in all three languages, thought the peak in Russian is less significant than in Latvian and English.

@@

% There is higher amount of Russian tweets, but Latvian and English tweet volume stays constant.

@@

We have seen that the reflection of an event on Twitter depends on the language. How are languages used individually? Do Twitter users tweet in a one language all the time, or switch between languages depending on the context?

@@

Another interesting accounts that tweet equally in Latvian and Russian, but do not tweet in English are the accounts of a library and a football club. It worth noting that mixing languages is not common in Latvian Internet. For example, large media portals tend to have separate Latvian and Russian web-sites and separate Twitter handlers.

To illustrate the language usage pattern between multilingual users, their first most frequently used language, their second most frequently used language and their third most frequently language were identified. Latvian is not only the most used language among the monoligual users, but also is the first and second most common choice between the multilingual users. Russian is the most frequent third choice, as Table~\ref{tab:language-use} shows.

@@

Regarding the social aspect of the corpus, the future work might have a look into why topics are language dependant? How separated/unified is the society? What are the controversial topics.

>>>>

\begin{abstract}
Twitter has become a rich source for linguistic data. Here, a possibility of building a trilingual Latvian-Russian-English corpus of tweets from Riga, Latvia is investigated. Such a corpus, once constructed, might be of great use for multiple purposes such as training machine translation models, examining cross-lingual phenomena and studying the population of Riga. This pilot study shows that it is feasible to build such a resource. The resulting corpus is made publicly available and can be used to construct a large comparable corpus.
\end{abstract}

@@

Comparable corpora are widely used by the natural language processing community to build machine translation or information retrieval models. The goal of this work is to investigate whether it is possible to build a comparable linguistic resource that originates from one specific location -- Riga, Latvia. Riga is a great location for this because it is a multilingual city in which Latvian and Russian are both widely used in everyday life, and English is used a lingua france in tourism and commerce. Being linked to a specific location, the resulting corpus might benefit not only the language engineers and linguists, but also social scientists.

@@

The analysis of the tweets, shows that they reflect real world events, such as national celebrations or international political affairs. While all three languages are represented---45.5\% tweets are in Latvian, 33.9\% in Russian and 20.7\% in English---not all events are equally discussed in all languages (Section~\ref{sec:timeline}). By studying users' tweeting habits, we see that the majority of users (83.3\%) mostly tweets in one language (Section~\ref{sec:lang-use}).

The properties of the corpus correspond to the expectation to reflect the real world events and language use proportion, but its size is too small to draw solid conclusions. However, the construction of a reliable corpus is the matter of the data collection procedure, because, as this study shows, there is relevant data on Twitter.

The resulting dataset is distributed as a supplement to the paper and available online at \texttt{http://...}

@@

251\,083 tweets were collected within the period from the 1st of November 2016 to the 31st of March 2017. On April 14th 2017, the collection was rehydrated by querying the Twitter API with the collected tweet IDs to get rid of the deleted tweets. In addition, the tweets that originated from retweets were added to the collection: the JSON representation of a retweet includes the original tweet, which was extracted and added to the collection. The rehydrated and expanded collection resulted in a total of 220\,883 tweets. Table~\ref{tab:tweet-counts} summarises the number of collected tweets.

@@

Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare.\footnote{\url{https://foursquare.com}} This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' most of the time follow a predefined template and thus do not reflect real language use. # What does this mean and why?

@@

Table~\ref{tab:client-counts} shows the top ten most popular clients in the rehydrated collection. Together with the tweets originating from Foursquare, tweets from Instagram,\footnote{\url{https://www.instagram.com}} an image sharing service, and Endomondo,\footnote{\url{https://www.endomondo.com}} a workout tracking service, were removed. Tweets written using the ``Word Cities'' client which posts weather reports and the Linkis client---a promotion website---were also removed.

The final collection resulted in 136\,067 tweets which are in Latvian, Russian or English and created after the 1st of November 2016. The language of a tweet is provided by the corresponding field in the tweet JSON representation.

@@

Figure~\ref{fig:timeline} shows the bandwidth of tweets over time for all three languages. There are several peaks in Twitter usage. Some of them affect all three languages, as in early January, some of them affect only one language, as in late January. # Any significance? Ah, I see next paragraph.

If the Twitter behaviour is affected by the events in the real world, then the peaks should correspond to events in the real word. The difference in peaks could then be explained as there are different real word events that trigger discussions on Twitter in Latvian, Russian and English. Table~\ref{tab:timeline-lang-corr} suggests, that tweets in Latvian and English share similar behaviour. The Russian tweet timeline is distinct from both timelines, tough its behaviour is more similar to the Latvian timeline than to English.

@@

11th of November is \href{https://en.wikipedia.org/wiki/L\%C4\%81\%C4\%8Dpl\%C4\%93sis_Day}{L\={a}\v{c}pl\={e}sis Day}, a memorial day for soldiers who fought for the independence of Latvia. 18 November is the Proclamation Day of the Republic of Latvia. Numerous events take place across the country including candle placing on and by the wall of Riga Castle, torchlight processions and fireworks. The volume of tweets peaks in all three languages, though the peak in Russian is less significant than in Latvian and English.

@@

% There is a higher amount of Russian tweets, but Latvian and English tweet volume stays constant.

@@

We have seen that the reflection of an event on Twitter depends on the language. How are languages used individually? Do Twitter users tweet in one language all the time, or switch between languages depending on the context?

@@

Another interesting accounts that tweet equally in Latvian and Russian, but do not tweet in English are the accounts of a library and a football club. It worth noting that mixing languages is not common in Latvian Internet. For example, large media portals tend to have separate Latvian and Russian web-sites and separate Twitter handles.

To illustrate the language usage pattern between multilingual users, their first most frequently used language, their second most frequently used language and their third most frequently language were identified. Latvian is not only the most used language among the monoligual users but also is the first and second most common choice between the multilingual users. Russian is the most frequent third choice, as Table~\ref{tab:language-use} shows.

@@

Regarding the social aspect of the corpus, the future work might have a look into why topics are language dependant? How separated/unified is the society? What are the controversial topics?
